Topic categorization in Portuguese news articles\\ \thanks{Fundação para a Ciência e Tecnologia}

### 2018-01-03 article.tex
References

The challenges of document classification have been well studied
within the machine learning research field of study. Given a corpus of
already classified documents, several algorithms might be applied to
train a classifier capable of determining the category of additional
articles.

@@

\section{Methods}
%TODO
we first needed to choose and obtain a suitable dataset, and then
cleaning and preparing it to be used to train the classifiers.

@@

%TODO
A document-term matrix was then calculated using the
\texttt{DocumentTermMatrix} method from the \texttt{tm} package.
For each of the 5600 documents preset in the train set the matrix
included all the terms which met the following requirements:

\begin{itemize}
    \item the term length was between 3 and 30 characters (to discard
        things like URLs or badly tokenized sentences)
    \item the term appeared in the document at least twice
    \item the term appeared at least in 10 documents
\end{itemize}
The algorithm used to weigh the terms in the matrix was the
\textit{tf-idf}\cite{christopher2008introduction}, and resulted in 3234 terms.

%TODO matrix stats(sparcity etc)
% inspect(dtm)

>>>>

The challenges of document classification have been well studied
within the machine learning research field of
study\cite{borko1963automatic,sebastiani2002machine,rubin2012statistical}.
Given a corpus of already classified documents, several algorithms
might be applied to train a classifier capable of determining the
category of additional articles.

@@

\section{Methods}
In order to train and evaluate classification algorithms, we first
needed to choose and obtain a suitable dataset. Preferably, this
dataset would contain documents were already categorized, which would
allows to skip the time and effor-consuming step of manually
categorizing the articles ourselfs. Once this dataset was chosen and
obtained, we would then clean and prepare it to be used to train the
classifiers.

@@

A document-term matrix was then calculated using the
\texttt{DocumentTermMatrix} method from the \texttt{tm} package.
For each of the 5600 documents preset in the train set the matrix
included all the terms which met the following requirements:

\begin{itemize}
    \item the term length was between 3 and 30 characters (to discard
        things like URLs or badly tokenized sentences)
    \item the term appeared in the document at least twice
    \item the term appeared at least in 10 documents
\end{itemize}
The algorithm used to weigh the terms in the matrix was the
\textit{tf-idf}\cite{christopher2008introduction}. The obtained matrix
presented a sparsity of 99\% and contained 3234 distinct terms.
